---
title: "TL:DR on FCC providers list"
date: last-modified
format: 
  html:
    code-fold: true
engine: knitr
---

```{r}
#| label: utility functions
source("R/table_with_options.R")
```

The data set was downloaded the 26-04-2024 adn was apparently avaiable since 11-03-2024. 

A csv version of it is available in `data/`

```{r}
#|label: read csv

csv_path <- "data/FCC_Providers_11_03_2024.csv"

col_classe <- c("Provider.Name" = "character", 
               "Affiliation" = "character",
               "Operation.Type" = "factor",
               "FRN" = "character",
               "Provider.ID" = "character")

isp <- read.csv(csv_path, colClasses = col_classe)

table_with_options(isp)
```

# Summary:

As we can see this table have 4 341 rows and 5 columns.

How many unique value do we have per column:

```{r}
#|label: count me some unique
sum_table <-  apply(isp, 2, function(x)length(unique(x))) |>
                    as.data.frame()
names(sum_table) <- c("Count of unique values")

knitr::kable(sum_table)
```

We can start by `Operation.Type` that can only take: "ILEC" or "Non-ILEC"

```{r} 
#| column: margin
#| label: Operation type
knitr::kable(table(isp$Operation.Type))
```

It seems `Affiliation` has the same number than `Provider.ID`. 

Provider.ID is close but a bit less number than `FRN` whcih seems to be the unique identifier of this table. 

