---
title: "FCC providers list"
date: last-modified
format: 
  html:
    code-fold: true
engine: knitr
---

```{r}
#| label: utility functions
source("R/table_with_options.R")
```

We have multiple sources from FCC to define a provider. 

- Two are coming from the Broadband Data Collection:

  * The first one from [FCC](https://us-fcc.app.box.com/v/bdcprovideridtable), accessed the 26-04-2024

  * The second is comming from our pipelines of NBM (June 2023 release)

- One is comming from the Emergency Broadband Benefit[^ebb] and can be downloaded [here](https://www.fcc.gov/sites/default/files/ebb.provider.list_.xlsx) (accessed the 26-04-2024)    

[^ebb]: The Emergency Broadband Benefit is the precursor of the Affordable Connectivity Program, source: [https://www.fcc.gov/broadbandbenefit](https://www.fcc.gov/broadbandbenefit)

::: {.callout-tip}
- [fcc.utils](https://github.com/ruralinnovation/fcc.utils) is a small R package that contains providers data and is used here

- Data is downlable using the download button
:::
 
# What does this data look like?

## Broadband Data Collection versions:

```{r}
#|label: read csv
library("fcc.utils")

isp <- fcc.utils::fcc_provider
```

`FCC provider` has 4 341 rows and 5 columns.

Those columns are: 

- `Provider.Name`: Same than Brand Name?

- `Affiliation`: Same number than `Provider.ID`

```{r} 
#| column: margin
#| label: Operation type

op_type <-as.data.frame(table(isp$Operation.Type, dnn = "Type"), responseName = "Nb." )

knitr::kable(op_type)
```

- `Operation.Type`: Only two options "ILEC" or "Non-ILEC"

- `FRN`: **F**CC **R**egistration **N**umber; "number of the entity that submited the data". It is supposed to be a string of 10 characters (with padding 0). Slighly more number than `Provider.Name` and seems to be the primary key.

- `Provider.ID`: An ID for `Affiliations`

```{r}
#|label: display ISP from BDC
table_with_options(isp)
```

# EDA / Analysis 

## How many unique values do we have per column:

```{r}
#| label: count me some unique
#| column: margin
#| tbl-cap: Coutnt of unique values / columns
sum_table <-  apply(isp, 2, function(x)length(unique(x))) |>
                    as.data.frame()
names(sum_table) <- c("Count of unique values")

knitr::kable(sum_table)
```

We can confirm that:

- `FRN` here is unique for every row in this data set (our primary key)

- We have a bit less `Provider.Name` (4321 / 4341) than `FRN`

- We have the same number of `Affiliations` and `Provider.ID`

- The number of `Provider.ID`/`Affiliations` is smaller than FRN. 

A quick check indicate that all `Provider.ID` are 6 characters.
`FRN` is also always 10 characters.

:::{.aside}
Hence the one with 7 characters in FCC NBM is probably an error.
:::

### What are the Provider Name thata re sharing multiple FRN:

```{r}
#| label: case where same name have different FRN
provider.name_by_FRN <- sapply(split(isp$FRN, isp$Provider.Name), function(x) unique(x))

multiple.frn <- provider.name_by_FRN[lengths(provider.name_by_FRN)> 1]

value = sapply(multiple.frn, toString) |> unlist()

provider.name_by_FRN.dat <- data.frame(Provide.Name = names(multiple.frn),
                                   FRN = sapply(multiple.frn, toString)
)

table_with_options(provider.name_by_FRN.dat)
```


### How FRN are split between "Affiliations"?

We probably have case where companies have same name ("Farmers Mutual Telephone company") but we have probably havfe company that have multiple FRN ("Grand Mound Cooperative Telephone Association").

```{r}
#| label: spliting number of affiliation by frn
#| column: margin
#| tbl-cap: Affiliation per FRN
FRN_by_affiliations <- sapply(split(isp$FRN, isp$Affiliation), function(x) length(unique(x)))
FRN_by_affiliations.dat <- data.frame(Affiliations = names(FRN_by_affiliations), 
                                      count_frn = FRN_by_affiliations)

cnt_affiliation <- as.data.frame(table(FRN_by_affiliations.dat$count_frn), responseName = "")
cnt_affiliation[["Num. of Affiliations / FRN"]] <- ifelse(as.numeric(cnt_affiliation[[1]]) < 10,  cnt_affiliation[[1]], "10+")

cnt_affiliation <- aggregate(cnt_affiliation[[2]], list(cnt_affiliation[["Num. of Affiliations / FRN"]]), sum)

names(cnt_affiliation) <- c("Number of Affiliations / FRN", "Count")


knitr::kable(cnt_affiliation[c(1, 3:9, 2 ),], row.names = FALSE)
```


 As expected most of of the relations FRN / provider are one to one (3135).

# Data set with mail address and phone numbers

This data set is one that is coming from the ACP program.

```{r}
#| label: address and phone
isp_contact <- read.csv("data/bb-provider_list.csv")
table_with_options(isp_contact)
```
