---
title: "TL:DR on FCC providers list"
date: last-modified
format: 
  html:
    code-fold: true
engine: knitr
---

```{r}
#| label: utility functions
source("R/table_with_options.R")
```

We have multiple sources from FCC to define a provider. 

- Two are coming from the Broadband Data Collection:

  * The first one from [FCC](https://us-fcc.app.box.com/v/bdcprovideridtable), accessed the 26-04-2024

  * The second is comming from our pipelines of NBM (June 23 release)

- One is comming from the Emergency Broadband Benefit[^ebb] and can be downloaded [here](https://www.fcc.gov/sites/default/files/ebb.provider.list_.xlsx) (accessed the 26-04-2024)    

[^ebb]: The Emergency Broadband Benefit is the successor of the Affordable Connectivity Program, source: [https://www.fcc.gov/broadbandbenefit](https://www.fcc.gov/broadbandbenefit)

::: {.callout-tip}
A csv version of each is available in `data/` and the data can be downloaded with the download button. 
:::
 
# What does this data look likes?

## Broadband Data Collection versions:

```{r}
#|label: read csv

csv_path <- "data/FCC_Providers_11_03_2024.csv"

col_classe <- c("Provider.Name" = "character", 
               "Affiliation" = "character",
               "Operation.Type" = "factor",
               "FRN" = "character",
               "Provider.ID" = "character")

isp <- read.csv(csv_path, colClasses = col_classe)
```

As we can see this table have 4 341 rows and 5 columns.

Those columns are: 

- `Provider.Name`: Sam than Brand Name?

- `Affiliation`: Same number than `Provider.ID`

```{r} 
#| column: margin
#| label: Operation type

op_type <-as.data.frame(table(isp$Operation.Type, dnn = "Type"), responseName = "Nb." )

knitr::kable(op_type)
```

- `Operation.Type`: Only two options "ILEC" or "Non-ILEC"

- `FRN`: **F**CC **R**egistration **N**umber; "number of the entity that submited the data". It is supposed to be a string of 10 characters (with padding 0). Slighly more number than `Provider.Name` and seems to be the primary key.

- `Provider.ID`: An ID for `Affiliations`

```{r}
#|label: display ISP from BDC
table_with_options(isp)
```

 
# Data set with mail address and phone number

```{r}
#| label: address and phone
isp_contact <- read.csv("data/bb-provider_list.csv")
table_with_options(isp_contact)
```


# EDA / Analysis 

How many unique value do we have per column:

```{r}
#|label: count me some unique
sum_table <-  apply(isp, 2, function(x)length(unique(x))) |>
                    as.data.frame()
names(sum_table) <- c("Count of unique values")

knitr::kable(sum_table)
```

A quick check indicate that all `Provider.ID` are 6 characters (hence the one with 7 in FCC NBM is probably an error). `FRN` is also always 10 characters.

### How FRN are split between Affiliations:

```{r}
FRN_by_affiliations <- sapply(split(isp$FRN, isp$Affiliation), function(x) length(unique(x)))
FRN_by_affiliations.dat <- data.frame(Affiliations = names(FRN_by_affiliations), 
                                      count_frn = FRN_by_affiliations)


knitr::kable(table(FRN_by_affiliations.dat$count_frn))
```


Most of it are one for one.