---
title: "EDA on ISP"
date: last-modified
format: 
  html:
    code-fold: true
engine: knitr
---

```{r}
#| label: utility functions
table_with_options <- function(x){DT::datatable(x, rownames = FALSE,
              extensions = 'Buttons', 
                options = list(
                          dom = 'Blfrtip',
                          buttons = list('copy', 'print', list(
                            extend = 'collection',
                            buttons = c('csv', 'excel'),
                            text = 'Download')
                                    )
                          )
            )}
```

We are starting a first exploratory data analysis around ISPs in the FCC BDC data set. It should be kept in mind that an ISP can be multiple time in the same location (offering multiple service). 

The query that generated this first pass at it is here: 

```SQL
SELECT 
    brand_name,
    state_abbr,
    technology,
    provider_id,
    count(*) as cnt_services
FROM
    staging.june23
GROUP BY 
    brand_name, provider_id, state_abbr, technology
ORDER BY 
    cnt_services DESC;
```

The name of the column match FCC [description](https://us-fcc.app.box.com/v/bdc-data-downloads-output). We are just adding `cnt_services` which is a count of **services**. In one locations you can have multiple services with different providers, technology and speeds provides (sometimes one providers can have multiple technology and/or multiple speeds). 

It can be explored here: 


```{r}
#| label: read and display ISP
isp <- read.csv("data/isp_v2.csv")
colnames(isp) <- c("brand_name", "state_abbr", "technology",
"provider_id", "cnt_services", "cnt_total_locations", "cnt_block_presence")
table_with_options(isp)
```


## Numbers for context: 

```{r}
#|  label: filter sat
filter_sat <- c(60, 61, 70)
isp_slim <- isp[! isp$technology %in% filter_sat, ]
```

In June 2023 we have `r length(unique(isp$brand_name))` unique ISP brand name but only `r length(unique(isp_slim$brand_name))` if we remove the ISP that only offer Satellite or Unlicensed Wireless. 

For the rest of the analysis I will not take into account Satellite data and Unlicensed Wireless.

Our first step will be to try having some "unique"  brand name so we can be confident we are correctly counting the same ISP (or not).  

## Organize a bit `brand_name` and `provider_id` 

 ```{r}
 #| label: some cleaning
isp_slim$brand_name <- tolower(isp_slim$brand_name)
 ```

It seems that we have:  
    - brand name with and without capital letter (VERIZON, Verizon): if we `tolower` brand name we get `r length(unique(isp_slim$brand_name))` unique brand name.

```{r}
#| label: agg per brand name
isp_agg <- aggregate(isp_slim["cnt_services"], isp_slim["brand_name"], sum)
table_with_options(isp_agg[order(isp_agg$cnt_services, decreasing = TRUE), ])
```

I have done a smaller `.csv` just with `brand_name` `provider_id` and `cnt_services` just to inspect what is the relation between them (1 to 1 / 1 to many). Outside of typos we should not have many to many relation. 

```sql
SELECT
    brand_name,
    provider_id,
    count(*) cnt_services
FROM staging.june23
GROUP BY brand_name, provider_id
ORDER BY cnt_services desc;
```

```{r}
isp_list <- read.csv("data/isp_prov.csv")
isp_list$ct <- 1  
isp_list$name_id <- ave(isp_list$ct, isp_list$provider_id, FUN = sum)
#View(isp_list[!is.na(isp_list$new_name),])
```

#### TCT

```{r}
#| label: TCT identify
table_with_options(isp_list[grepl("^TCT ", isp_list$brand_name) ,])
isp_list[isp_list$provider_id == 410172,]
```

TCT has some non-conventional names but nearly all of them has the same `provider_id`. An other "TCT" exist but with a different provider_id (`410172`) shared with two other brand name. I will assume that all of this `TCT XXX` are the same and provide them with a temporary name `TCT_131366`.

```{r}
#| label: TCT update
isp_list$new_name[grepl("^TCT ", isp_list$brand_name)] <- "TCT_131366"
```

#### Windstream 

Windtream present a similar case but the position of Windstream is not always the first word (` Georgia Windstream, LLC`). I went with the solution than TCT: `Windstream_131413`


```{r}
#| label: windstream identify and update
table_with_options(isp_list[grepl("Windstream", isp_list$brand_name) ,]) 

isp_list$new_name[grepl("Windstream", isp_list$brand_name)] <- "Windstream_131413" 
```

#### Acentek/Acentek

It exists in both forms (`tolower()` will correct it) but it is also sharing it's `provider_id` with some non-conventional "name": 

```{r}
#| label: acentek
table_with_options(isp_list[isp_list$provider_id == 130008,])
isp_list$new_name[isp_list$provider_id == 130008] <- "acentek"
```

For now I will go with attributing them to Acentek but an other option will be to just remove them. 

#### Mediacom - Bolt

```{r}
#| label: Mediacom bolt
table_with_options(isp_list[grepl("Mediacom|Bolt", isp_list$brand_name) ,]) 
isp_list$new_name[isp_list$provider_id == 130804] <- "mediacom_bolt"
```

It appears that Bolt and Mediacom share the same `provider_id` and are together in some `brand_name`. I think we should regroup them but this definietly more domain knowledge than the one I have!

#### Comporium 

```{r}
#| label: comporium
table_with_options(isp_list[isp_list$provider_id == 131125,])
isp_list$new_name[isp_list$provider_id == 131125] <- "comporium"
```

For this one I am for renaming them "comporium"

#### Armstrong

```{r}
#| label: Armstrong
table_with_options(isp_list[isp_list$provider_id == 130071,])
isp_list$new_name[isp_list$provider_id == 130071] <- "armstrong"
```

Idem label to "armstrong"?

#### TEC 

```{r}
#| label: TEC
table_with_options(isp_list[isp_list$provider_id == 131311,])
isp_list$new_name[isp_list$provider_id == 131311] <- "TEC"
```

#### PUD 

```{r}
#| label: PUD
table_with_options(isp_list[isp_list$provider_id == 290075,])
isp_list$new_name[isp_list$provider_id == 290075] <- "PUD"
```

Label to "PUD" ?

#### GoNetspeed?

I am unsure about that one. We can regroup the two `GoNetspeed` but we are lacking information for the rest. 

```{r}
#| label: gotnetspeed
table_with_options(isp_list[isp_list$provider_id == 131378,])
```

### Multiple brand_name with same `provider_id`

- provider_id: 131413, same provider id than windstream but probably a different entity?

- provider_id 130485: look very similar (`SHLB` in the name 8/10)

- list of `provider_id` associated with multiple brand name: 

    * 130074

    * 190233

    * 160127

    * 150277

    * 150266

    * 130183

    * 300192

    * 131362

    * 130877

    * 130757 (regroup Long lines?)



## TODO list:

[ ] provider_id: 586211


