---
title: "EDA on ISP"
date: last-modified
format: 
  html:
    code-fold: true
engine: knitr
---

```{r}
#| label: utility functions
table_with_options <- function(x){DT::datatable(x, rownames = FALSE,
              extensions = 'Buttons', 
                options = list(
                          dom = 'Blfrtip',
                          buttons = list('copy', 'print', list(
                            extend = 'collection',
                            buttons = c('csv', 'excel'),
                            text = 'Download')
                                    )
                          )
            )}
```

We are starting a first exploratory data analysis around ISPs in the FCC NBM data set. It should be kept in mind that an ISP can be multiple time in the same location (offering multiple service).  

The query that generated this first pass at it is here: 

```SQL
SELECT 
    brand_name,
    state_abbr,
    technology,
    provider_id,
    count(*) as cnt_services
FROM
    staging.june23
GROUP BY 
    brand_name, provider_id, state_abbr, technology
ORDER BY 
    cnt_services DESC;
```

It can be explored here: 


```{r}
#| label: read and display ISP
isp <- read.csv("data/isp.csv")
table_with_options(isp)
```


## Numbers for context: 

```{r}
#|  label: filter sat
filter_sat <- c(60, 61, 70)
isp_slim <- isp[! isp$technology %in% filter_sat, ]
```

In June 2023 we have `r length(unique(isp$brand_name))` unique ISP brand name but only `r length(unique(isp_slim$brand_name))` if we remove the ISP that only offer Satellite or Unlicensed Wireless. 

For the rest of the analysis I will not take into account Satellite data and Unlicensed Wireless.


```{r}
#| label: agg per brand name
isp_agg <- aggregate(isp_slim["cnt_services"], isp_slim["brand_name"], sum)
head(isp_agg[order(isp_agg$cnt_services, decreasing = TRUE), ], 10)
```

 ```{r}
isp_slim$brand_name <- tolower(isp_slim$brand_name)
#length(unique(isp_slim$brand_name))
 ```

It seems that we have:  
    - brand name with and without capital letter (VERIZON, Verizon): if we `tolower` brand name we get `r length(unique(isp_slim$brand_name))` unique brand name.

